### 2019-09-12 manuscript/manuscript_removed_text.Rmd
Update manuscript_removed_text.Rmd

---
title: "manuscript_removed_text"
output: pdf_document
---

Globally, forests play an important role in the carbon cycle and are a major component of global carbon dioxide budgets [@luyssaert_old-growth_2008]. They show higher levels of productivity than non-forest terrestrial ecosystems [@del_grosso_global_2008], and as a result achieve significant carbon sequestration and storage. Estimating the total role of forests in the carbon cycle is challenging, but studies indicate that old growth forests alone sequester up to 1.4 GtCyr^-1^ [@malhi_carbon_1999], while the total sequestration of carbon by established forests globally could be up to 2.4 GtCyr^-1^, with the largest sinks being in old-growth tropical forests [@pan_large_2011]. As atmospheric carbon dioxide levels continue to rise, with consequences for global climate, there is increasing recognition that proper protection and management of forest resources will have an important role to play in mitigating climate change. Understanding the patterns of forest productivity on a global scale, and the drivers behind them, is therefore a priority in forest research.

There are two major questions to understand: firstly, how forest productivity varies globally - and specifically which areas show the greatest peaks in productivity -; and secondly, which climate variables drive this variation. On a global scale, the productivity of forests varies with latitude, with a general trend of increasing productivity towards the tropics [@beer_terrestrial_2010; @jung_global_2011]; however the exact nature of this pattern, and how it varies by component of productivity, is poorly understood. This latitudinal gradient is most likely to be explained by climatic gradients in temperature, precipitation, length of growing season, and combinations of the above. Productivity is influenced by a range of climatic drivers, including mean annual temperature (MAT) and mean annual precipitation (MAP) [@del_grosso_global_2008], but doesn't necessarily respond linearly to these drivers. Disentangling the shape of productivity responses to climate drivers will enable better predictions of future responses under climate change.

*The components of forest productivity.* During photosynthesis, plants capture carbon dioxide from the atmosphere. The gross primary productivity (GPP) of an ecosystem is the gross uptake, via photosynthesis, of carbon dioxide by plants in that ecosystem. Only a fraction of the carbon captured is assimilated into plant tissue; the rest is used in autotrophic respiration (R~auto~). The component of GPP that is stored as plant material is the net primary productivity (NPP) of an ecosystem. Net primary productivity can therefore be expressed as:

$$ NPP = GPP + R~auto~ $$

Currently, GPP cannot be measured directly by observing total ecosystem photosynthesis. Instead field estimates of GPP have to be derived based on modelling and extrapolation of eddy-covariance studies and measurements of net ecosystem exchange (NEE) [@clark_reviews_2017].

In contrast, NPP can be calculated through direct field observations. In order to achieve greater accuracy in estimating NPP, NPP is often broken down into its component parts, with aboveground NPP (ANPP) and belowground NPP (BNPP) considered separately to each other. The components included when estimating ANPP and BNPP often vary between studies, depending on the intensity of fieldwork effort. At its most basic level, ANPP can be expressed as:

$$ ANPP = NPP~stem~ + NPP~branch~ + NPP~canopy~ $$
where NPP~stem~ is the annual woody increment of all stems above a specified diameter at breast height (DBH), NPP~branch~ is annual woody branch turnover, and NPP~canopy~ is annual foliage production, including leaves, twigs, and reproductive structures. ANPP may also include NPP~VOC~, the annual emission of volatile organic compounds, and NPP~herbivory~, the annual consumption of plant matter by herbivores, but these components are often excluded from field observations as they are much harder to quantify. Other components of aboveground productivity that remain largely unquantified include epiphytes, hemiepiphytes, and understory plants [@clark_reviews_2017]. The majority of current ANPP estimates are based on the assumption that the contribution of these components to overall NPP is insignificant.

There are two major subcomponents of BNPP, which can be expressed as:

$$ BNPP = NPP~coarse root~ + NPP~fine root~ $$
where NPP~coarse root~ is the annual production of coarse roots (typically roots >2mm diameter), and NPP~fine root~ is the annual production of fine roots (typically roots <2mm diameter) [@aragao_above-_2009]. Calculations of BNPP may also include NPP~exudation~, a measure of annual carbon losses through root exudation, and NPP~symbionts~, the annual carbon allocation to mycorrhizae and legumes, but, as before, this is challenging to quantify and is often excluded from field observations.

BNPP is a poorly understood component of total ecosystem productivity, primarily because of the challenges in gaining accurate field measurements. Coarse root productivity is often estimated via extrapolation of NPP~stem~ estimates using allometries that may not have been empirically verified [@clark_net_2001]. NPP~fine root~ is easier to quantify through soil cores and minirhizotrons, however, sampling tends to be limited to the surface soils, with very few studies sampling to depths below 3 metres [cite]. As a result, it is possible that BNPP is currently significantly underestimated, despite being a hugely significant component of total ecosystem productivity [@pan_large_2011].

*Which factors influence primary productivity?* Primary productivity is influenced by many factors, which often act across a range of scales, and may show interactive effects with each other. On a local scale, stand age [@litton_carbon_2007; @gillman_latitude_2015], management [@simova_enigma_2017]; nutrient availability [@aragao_above-_2009]; and altitude [@girardin_net_2010; @malhi_variation_2017] all impact forest productivity. On a global scale, changes in primary productivity are influenced by climatic variables and abiotic gradients, such as the length of growing season [@michaletz_convergence_2014]. There is some debate over the precise relationship between these drivers and productivity; while MAT and MAP have been argued to be significant predictors of productivity [@chu_does_2016], other studies have found that the correlation between productivity and MAT is a factor of the relationship between productivity and growing season length [@kerkhoff_plant_2005; @malhi_productivity_2012; @michaletz_convergence_2014; @michaletz_drivers_2018]. Improving understanding of how these factors interact to control global patterns in primary productivity is essential to understanding the global carbon cycle.

Current research into how primary productivity varies with latitude is inconclusive, and - though it has primarily focussed on patterns of GPP, NPP, and ANPP - indicates that different components of productivity may show different relationships with latitude. Gross primary productivity is generally thought to be highest in the tropics. Modelling of global terrestrial ecosystem GPP though upscaling and calibration of eddy flux measurements indicates a peak in GPP in the tropics, with the highest levels in tropical forests [@beer_terrestrial_2010; @jung_global_2011]. This is corroborated by analysis of site-level GPP measurements, which show a strong positive correlation between GPP and MAT and MAP [@luyssaert_co2_2007], with the highest GPP values reported in tropical forests. The influence of latitude on global patterns of NPP is less clear than that of GPP. Simova and Storch (2017) found that, as with GPP, NPP decreases with latitude, peaking in the tropics. However, other studies have found the highest values of NPP in temperate forests [@luyssaert_co2_2007; @huston_global_2009]. Because of the challenges in accurately obtaining unbiased measures of belowground productivity, many studies focus on ANPP in preference to measures of NPP. Studies on global patterns of ANPP are equally inconclusive: Gillman et al. (2015) found a weak negative relationship between ANPP and latitude, with the relationship becoming stronger in older forest stands. These findings were echoed in other studies, which have found weak or no relationships between ANPP and latitude [@huston_global_2009].

Furthermore, there is evidence that different components of productivity show individual responses to climate drivers. For example, increases in GPP have been reported to saturate above 25$^\circ$C MAT [@larjavaara_temperature_2012], while increases in NPP are recorded to saturate above 10$^\circ$C MAT [@luyssaert_co2_2007]. Similarly, allocation to different components of ANPP varies with climate. Within the tropics, allocation to canopy NPP appears fairly consistent, with significantly greater variation in allocation to woody and belowground NPP [@litton_carbon_2007; @malhi_allocation_2011]. Allocation to these structural biomass components has been shown to increase with water availability [@litton_carbon_2007; @bloom_decadal_2016], and is highest in the wet tropics, indicating that control of woody productivity by MAP may be more significant than control of foliar productivity. However, these studies are regional, meaning that our understanding of variation in allocation and its relationship to climate on a global scale remains limited. In addition, allocation is also influenced by stand age [@de_lucia_forest_2007], nutrient availability [@litton_carbon_2007] and forest structure [@taylor_greater_2019], which can make it challenging to disentangle the effects of climate.

Data that control for stand age and standardize methodologies are required to resolve this question. Here, we use a comprehensive global database of forest carbon fluxes to explore global patterns in productivity. We explore five broad questions, described in greater detail in table 1:  
  1. How do carbon fluxes vary with latitude?  
  2. How do carbon fluxes relate to MAT and MAP?  
  3. How do carbon fluxes relate to other annual climate variables?  
  4. What is the role of seasonality in explaining variation in carbon fluxes?  
  5. Per month of the growing season, how does productivity vary with climate?

1. **Which climatic variables are the most important drivers behind the latitudinal pattern in primary productivity?** To date, the majority of studies have focussed on productivity responses to precipitation and temperature, while the influence of other climate variables on productivity remains under-explored. We utilise global climate datasets to investigate the relationships between productivity and a range of climate variables.  
2. **Do the components of primary productivity show variation in their responses to these climatic drivers?** Allocation to components of primary productivity appears to show some variation across climate and latitudinal gradients. We use global datasets of a range of primary productivity components to explore how significant this variation in allocation is.  
3. **Does climate explain the same proportion of variation in different components of primary productivity?** Review of current research shows that, while the relationship between GPP and latitude is consistent, NPP and ANPP show weaker relationships with latitude. This may be a result of varying sensitivities to climate drivers. We use mixed models to estimate the proportion of variation explained by climate for different components of productivity, to investigate how the relative importance of climatic drivers varies across components.

*Which climatic variables are the most important drivers behind the latitudinal pattern in primary productivity?* All major carbon fluxes show similar responses to climate variables. Of the climate variables tested, annual wet days, aridity, cloud cover, mean diurnal temperature range and precipitation seasonality were poor explainers of variation in primary productivity, explaining less than 20% of the variation in each of the carbon fluxes.
We found that both measures of mean annual precipitation used were significant (p<0.05) but poor predictors of productivity, explaining, with the exception of R~auto~, at most 37% of variation in carbon flux.

Looking across all productivity components, the climate variables which explain the highest amount of variation in carbon fluxes are temperature seasonality, mean annual temperature, annual temperature range, and annual frost days. Vapour pressure deficit came out as a strong predictor for ANPP~woody stem~, and potential evapotranspiration was a good predictor for BNPP~root~ and ANPP~woody stem~.

```{r echo=FALSE, out.width='100%', fig.cap = "Plots of primary productivity $(MgC~ha^-1~yr^-1)$ regressed against the climate variables identified as the best predictors: mean annual temperature; temperature seasonality; annual temperature range; and vapour pressure. Lines of best fit are plotted according to the best model selected during analysis. All regressions are significant $(p<0.05)$.", fig.pos='H'}
  knitr::include_graphics("combined_plots.png")
```

**Mean annual temperature.** All carbon fluxes show a linear increase with temperature.  
**Temperature seasonality and annual temperature range.** These two climate variables are strongly correlated, and so only temperature seasonality is presented in figure 4. Productivity is highest where temperature seasonality = 0, and at an annual temperature range of 15$^\circ$C or lower. GPP, NPP, ANPP, and R~root~ show a polynomial relationship with both variables, and BNPP~root~ shows a polynomial relationship with annual temperature range. As seasonality increases, these variables show a more rapid decrease in productivity. ANPP~foliage~ and ANPP~woody stem~ show a linear relationship with both variables.  
**Mean annual precipitation.** Despite being a poor predictor of productivity, the results indicate that, for the majority of fluxes, productivity increases with precipitation, up until a saturation point at between 3000 and 4000mm annual precipitation, above which productivity starts to decrease. The notable exception to this is GPP: the model indicates that GPP continues to increase with precipitation up to measures of at least 5000mm annually (p<0.0001, R^2^ = 0.33. Data above this point is not available, but the model trend indicates that the saturation point for this model is around 5000mm mean annual precipitation.  
**Potential evapotranspiration.** All fluxes, with the exception of BNPP~fine root~ and R~root~ show a strong increase with potential evapotranspiration at values below 1000mm. Between 1200 and 1700mm productivity peaks, and begins to decrease at values above 1800mm.

There was a significant interactive effect between MAT and MAP for GPP, BNPP~root~, BNPP~fine root~, ANPP, ANPP~woody stem~, and R~root~. There was a significant additive effect for R~auto~. NPP and ANPP~foliage~ showed no significant interactive or additive effect: including MAP as a second variable did not improve on the model including only MAT.

For the variables which showed a significant interactive or additive effect between MAT and MAP, no other climate variable, in combination with MAT, significantly improved on that model. For NPP, there was a significant interactive effect between MAT and water stress months, with this model explaining nearly 5% more variation in NPP than MAT alone. However, for ANPP~foliage~, no multivariate model improved on the univariate model including only MAT.

```{r echo=FALSE, out.width='100%', fig.cap = "Plots of primary productivity $(MgC~ha^-1~yr^-1)$ regressed against mean annual precipitation. Points are grouped into bins of 0 - 1000, 1001 - 2000, 2001 - 3000, and >3000mm mean annual precipitation, and lines of best fit plotted for mean annual precipitation values of 500, 1500, 2500, and 3500mm. All regressions are significant $(p<0.05)$.", fig.pos='H'}
  knitr::include_graphics("mat_map_interaction.png")
```

*Do carbon fluxes vary in their responses to climate, particularly with regards to allocation between fluxes?* Of all the flux pairings tested, the only ones to show a significant relationship with latitude were ANPP:NPP (p<0.01, R^2^ = 0.44) and ANPP:BNPP (p<0.001, R^2^ = 0.33). In both cases, allocation to ANPP increased with latitude, with the lowest allocation to ANPP in the tropical regions. BNPP:NPP decreased with latitude, though this relationship was non-significant.

| Database | Variables downloaded | Citation |
|-----------------------|--------------------------------------------|---------------------|
| WorldClim | Mean annual temperature; temperature seasonality; annual temperature range; mean annual precipitation | [@hijmans_very_2005] |
| WorldClim2 | Solar radiation | [@fick_worldclim_2017] |
| Climate Research Unit (CRU) time-series dataset v 4.03 | Cloud cover; annual frost days; annual wet days; potential evapotranspiration | [@harris_updated_2014] |
| Global Aridity Index and Potential Evapotranspiration Climate Database | Aridity; potential evapotranspiration | [@trabucco_global_2018] |
| TerraClimate | Vapour pressure deficit | [@abatzoglou_terraclimate_2018] |

Table: Sources of climate data

| Variable | Latitude | Mean annual temperature | Temperature seasonality | Potential evapotranspiration | MAT*MAP |
|----------|----------|----------|----------|--------------|----------|
| R~auto~ | 0.65 | 0.77 | 0.62 | 0.63 | NA |
| GPP | 0.64 | 0.63 | 0.72 | 0.36 | 0.64 |
| NPP | 0.50 | 0.52 | 0.52 | 0.33 | 0.56 |
| ANPP | 0.45 | 0.46 | 0.42 | 0.28 | 0.48 |
| ANPP~foliage~ | 0.54 | 0.63 | 0.54 | 0.35 | 0.64 |
| ANPP~woody~ ~stem~ | 0.17 | 0.23 | 0.12 | 0.19 | 0.28 |
| BNPP~root~ | 0.33 | 0.26 | 0.23 | 0.36 | 0.35 |
| BNPP~fine~ ~root~ | 0.15 | 0.11 | 0.16 | 0.11 | 0.19 |
| R~root~ | 0.22 | 0.21 | 0.30 | 0.01* | 0.28 |

Table: R^2^ values presented for latitude, three climate variables, and the best multivariate model for each carbon flux. Values marked * are not significant.

In this analysis we used a comprehensive global database, containing an unprecedented amount of data and representing all global forest ecosystems and all significant forest carbon fluxes, to comprehensively explore the relationships between climate and productivity on a global scale. Many of our findings support and clarify previously published work; however we used a much larger and more complete database than has previously been available, allowing for higher quality control and standardisation of data. This enabled us to control for factors not previously controlled for - such as stand age, methodology, flux components, and disturbance regime - to gain a stronger understanding of latitudinal and climatic effects on productivity.

Our results show that productivity decreases linearly with latitude. Climate explains a significant proportion of variation in all carbon fluxes, with temperature variables being the best predictors of productivity on this global scale. While other climate variables are significant predictors of productivity, none of them improve on the explanatory power of temperature-related variables in general or MAT specifically. Water availability is an important factor in explaining productivity on a global scale: we find a positive influence of precipitation at low MAP, with saturation at higher levels of MAP. There is a significant interaction between MAT and MAP.  We note that climate tends to explain a higher amount of variation in the major fluxes, with the amount of variation explained decreasing in smaller component fluxes.

We find that seasonality is an important factor in understanding patterns in productivity on a global scale: temperature seasonality and growing season length are strong predictors of productivity, though growing season length doesn't improve on MAT as a predictor. Within the growing season we find the influence of climate on productivity is smaller, but still significant for a number of carbon fluxes.

Despite the high standard of data quality in the ForC database, there are still significant limitations on our analyses. We standardised methodologies as far as possible, ensuring that there was consistency in the components included for each carbon flux. However, field techniques for estimating components of primary productivity are variable in their accuracy. This is particularly true for measures of BNPP, which are often estimated by extrapolating measures of ANPP. Even where BNPP is calculated through direct soil sampling, measures are often taken to insufficient depth, leading to underestimation of the flux. It is challenging to account for this type of variability in quality of field data, and our analyses largely assume good faith in data quality. While we have strong confidence in our analyses of well-understood fluxes, such as GPP, ANPP, and ANPP~woody~ ~stem~, we recommend caution in drawing firm conclusions from our analyses of BNPP, BNPP~fine~ ~root~, and, to some extent, NPP. We expect that, as field techniques improve the accuracy of estimates of these fluxes in the future, our understanding of the relationship of BNPP with climate variables will change. It is possible that the low R^2^ values for BNPP that we find are a result of lower accuracy in field measures of these fluxes, and that with improve standardisation of methods, we would find climate explains a higher proportion of variation.

Secondly, we emphasise that this study considers the relationship between productivity and climate on a coarse scale. Because of the nature of the ForC database, we were unable to access site-level MAT and MAP data for a number of sites, and lacked site-level data for other climate variables for all sites. As a result the climate data we used was primarily multi-year averages of annual climate measures accessed from global databases. While the quality of these data is high, we cannot expect that it captures detailed local variation in climate, and in particular will be unable to capture microclimate variation, which may have a strong influence on stand-level productivity. Although we find that MAT is the strongest predictor for climate, part of this effect may be a result of the higher-quality stand-level dataset available for MAT. With high-quality stand-level data for other climate variables we expect that their predictive strength would improve.

Furthermore, we expect that temporal resolution of climate data affects our results. Productivity is known to vary on a seasonal and annual scale in response to variation in climate. Our analysis relies on long-term averages in both climate variables and forest productivity estimates, and as a result averages out a large proportion of this variation. Therefore we recognise that the strong short-term effects of climate on productivity are likely to be masked. The limitations of low temporal resolution are highlighted in our analysis of growing season length. We note that our estimate of growing season length is imprecise, being based on monthly temperature, precipitation, and evapotranspiration averages. With accurate stand-level measurements of growing season length, based on daily, not monthly, climate measures, we would expect to record a stronger correlation between growing season length and productivity.

Although we recognise the inherent limitations present in this analysis, by focusing on mature, undisturbed forests, and by clarifying the components of NPP included in analysis, our study offers a clearer picture of how carbon cycling varies with latitude and climate.

Past studies have differed in their conclusions regarding the relationship between productivity and latitude. Our findings indicate that, for all carbon fluxes studied, productivity peaks in the tropical regions. There is strong evidence from our analyses that this increase in productivity towards the tropical regions is driven primarily by temperature-related effects. We find that temperature-related climate variables explain the highest proportion of variability in productivity. In contrast, we find that MAP is not a significant predictor of productivity, but does show a strong interaction with temperature. This indicates that the main mechanism through with precipitation influences productivity is in a secondary capacity, with its overall influence being mediated by the effects of temperature. This also suggests that although temperature explains a large proportion of variation, it cannot fully explain productivity alone, and a measure of water availability is necessary to fully understand patterns of productivity. This is likely to be significant in the future, as it suggests that the effects of a warming climate will be mediated through changes in precipitation regime.

There is debate over whether productivity in regions with higher MAT is higher because productivity inherently increases with air temperature during growing season months, or because higher MAT is correlated with longer growing seasons. Our analysis suggests that length of growing season is a key factor in understanding latitudinal patterns in productivity. We find that length of growing season is strongly correlated with productivity. This is supported by the strong correlation between temperature seasonality and productivity, indicating that aseasonality and year-round growing seasons have a strong influence over productivity.

These conclusions are supported by our analyses of the influence of climate within the growing season. We find much weaker correlations between climate variables and productivity when only growing season months are considered. There is evidence for a positive correlation between productivity and temperature for ANPP and ANPP~foliage~, suggesting that there is still a small positive effect of air temperature on productivity independent of growing season, however this effect is not significant for the remaining fluxes. This suggests that the primary positive effect of temperature on productivity is through the lengthening of the growing season in regions with higher MAT. However, we note that length of growing season is a less good predictor of productivity than MAT. As discussed above, this may be an artefact of imprecise estimation of length of growing season, however it does indicate that there are aspects of productivity explained by MAT which length of growing season alone cannot explain.

By including multiple fluxes in our analysis, we are able to explore the relationships between fluxes and their responses to climate. We find that fluxes are broadly consistent in their responses to climate drivers, suggesting that the importance of climate drivers on productivity is relatively invariant across fluxes. In addition, we show that carbon allocation between subsidiary fluxes does not vary significantly across broad climatic gradients. There is evidence that allocation between carbon fluxes does vary with factors such as stand age [@litton_carbon_2007], nutrient availability [@litton_carbon_2007; @gill_belowground_2016], forest structure [@taylor_greater_2019], elevation [@moser_elevation_2011], and water availability [@newman_above-_2006]. That we do not see consistent patterns of variation in allocation across climatic gradients indicates that the pressures influencing shifts in allocation between fluxes occur on local scales, and that shifts in allocation occur as a response to fluctuations in local environmental conditions.

Related to this, we find that climate more strongly explains variation in the major fluxes, with lower levels of variation explained in subsidiary fluxes. Although this could in part be related to a lack of standardisation in methodology across studies, as discussed above, it could also indicate that factors other than climate have significant influence over subsidiary fluxes. This is consistent with the results we find for allocation between fluxes: although climate has a significant influence on subsidiary fluxes, it is not the only factor that is important, and many other local- and regional-scale factors  influence the way carbon is allocated between subsidiary fluxes. We would expect that this would reduce the strength of the effect of climate on subsidiary fluxes.

**[needs strong concluding paragraph]**