### 2020-01-20 manuscript/manuscript-draft.tex
Table 2

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Global patterns of forest autotrophic carbon fluxes},
            pdfauthor={Rebecca Banbury Morgan, Valentine Herrmann, Norbert Kunert, Ben Bond-Lamberty, Helene C. Muller-Landau, Kristina J. Anderson-Teixeira},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{natbib}
\bibliographystyle{apalike}
\usepackage{longtable,booktabs}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\providecommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{Global patterns of forest autotrophic carbon fluxes}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{Rebecca Banbury Morgan, Valentine Herrmann, Norbert Kunert, Ben
Bond-Lamberty, Helene C. Muller-Landau, Kristina J. Anderson-Teixeira}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
    \date{}
    \predate{}\postdate{}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\usepackage{float} \usepackage{caption} \captionsetup[table]{font=footnotesize} \captionsetup[figure]{font=footnotesize}

\begin{document}
\maketitle

\subsubsection{Abstract}\label{abstract}

{[}\textbf{very rough start:}{]} \ldots{}.. Here, we drew upon \#
records from the Global Forest Carbon Database (ForC), representing all
major forest types and the nine most significant forest autotrophic
carbon flux (FACF) variables, to comprehensively explore how C cycling
in mature, undisturbed forests varies with latitude and climate on a
global scale. We show that, across all FACF variables analyzed, C
cycling decreases linearly with latitude --a finding that confirms
multiple previous studies and contradicts the idea that productivity of
temperate forests rivals that of tropical forests. The FACF variables
increase in proportion to one another, with no differences in allocation
detectable at this global scale, but differed in that latitude explained
a lower proportion of variation among subsidiary fluxes (in particular,
woody aboveground productivity and belowground net primary productivity,
\(BNPP\)). Climate explained a significant proportion (\#-\#\%) of
variation in all C fluxes analyzed (less for subsidiary fluxes), with
temperature variables in general and mean annual temperature (\(MAT\))
in particular being the best predictors of FACF on this global scale.
While other climate variables (\emph{e.g.}, \textbf{XX}) displayed
significant correlation with FACF, none of them had significantly better
explanatory power than \(MAT\). The effects of temperature were modified
by moisture availability, with reduced FACF under hot and dry conditions
and sometimes under very high precipitaiton (especially for \(BNPP\)).
FACF declined with temperature seasonality, but growing season length
doesn't improve upon MAT as a predictor. Within the growing season, the
influence of climate on C cycling is smaller but still significant for a
number of carbon fluxes. These findings clarify the big picture of how
FACF varies with latitude and climate on a global scale. As we enter a
period of accelerating climatic change, understanding of the fundamental
climatic controls on FACF sets a foundation for understanding patterns
of change.

\subsubsection{Introduction}\label{introduction}

(\textbf{revisit these first two paragraphs.})

\textbf{Forests play a critical role in regulating atmospheric
CO\textsubscript{2} and climate} (\textbf{Bonan}). Approximately
\textbf{\#\#} Gt of CO\textsubscript{2} cycle through Earth's forests
each year \textbf{REF-KAT}), and their net C sequestration
(\textasciitilde{}2.4 GtCyr\textsuperscript{-1}) offsets roughly 30\% of
anthropogenic fossil fuel emissions \citep{pan_large_2011}. As
atmospheric carbon dioxide levels continue to rise, driving climate
change, there is increasing recognition that protection and management
of forest resources will play an important role in mitigating climate
change (** REFS- e.g., IPCC 1.5C and land reports**). Carbon fixation,
allocation, and metabolism by trees set the basis for energy and
material flows in forest ecosystems and define their interactions with
Earth's changing climate.

\textbf{Forests' autotrophic carbon flux (FACF, bbl:define) influences
all organic matter stocks in ecosystems and is linked to cycling of
energy, water, and nutrients} (\textbf{REFS; \emph{Piao et al. 2010.
Forest annual carbon cost: a global‐scale analysis of autotrophic
respiration. Ecology 91(3), 652-661} }). Forest productivity sets the
energy ultimately available to heterotrophic organisms (including
microbes), in turn influencing their abundance (\textbf{REFS}) and
possibly diversity
(\href{https://www.annualreviews.org/doi/abs/10.1146/annurev.ecolsys.30.1.257}{Waide
et al. 2009}\textbf{REFS}). However, our ability to draw general
macroscopic conclusions regarding global variation in multiple
productivity variables with respect to climate has been limited in that
these analyses often mix forests that vary in stand age, disturbance
history, and/or management status; do not always sufficiently parse
related variables (\emph{e.g.}, combining net primary productivity
records with and without belowground components); and typically consider
only one or a few variables at a time.

\textbf{The productivity of forests varies with latitude}, showing a
general trend of decreasing productivity with latitude
\citep{beer_terrestrial_2010, jung_global_2011}. Studies agree that
productivity is lowest in the boreal regions, and increases into the
temperate regions
\citep{luyssaert_co_2007, huston_global_2009, beer_terrestrial_2010, jung_global_2011}.
However, evidence is inconclusive on whether productivity continues to
increase into the tropics, or whether it plateaus in temperate regions.
Evidence for this is further complicated by the fact that different
studies use different measures of productivity to explore these
relationships. For example, modelling of global terrestrial ecosystem
gross primary productivity (GPP) through upscaling and calibration of
eddy flux measurements indicates that GPP peaks in tropical forests
\citep{beer_terrestrial_2010, jung_global_2011};\href{https://onlinelibrary.wiley.com/doi/epdf/10.1111/gcb.14729}{Bagdley
et al. 2019}). This is corroborated by analysis of site-level GPP
measurements, which appear to reach their highest levels in tropical
forests \citep{luyssaert_co_2007}. \{\textbf{condense the previous two
sentences into one-- both primarily eddy flux}\} In contrast, some
studies suggest that the highest values of net primary productivity
(NPP) may be found in temperate forests
\citep{luyssaert_co_2007, huston_global_2009}, while others find NPP
highest in the tropics and decreasing with latitude
\citep{simova_enigma_2017}. Other studies have chosen to focus
exclusively on above-ground net primary productivity (ANPP), finding
evidence of a weak negative relationship between ANPP and latitude
\citep{huston_global_2009, gillman_latitude_2015}.

\textbf{Climate is a significant driver of productivity across broad
spatial scales} \citep{cleveland_relationships_2011}. The majority of
studies have focused on exploring the relationships of productivity to
mean annual temperature (MAT) and mean annual precipitation (MAP), as
the most commonly reported site-level climate variables. These variables
have the advantage that they describe broad trends in temperature and
water availability, and therefore capture a lot of global-scale
variation in climate. There is strong evidence that both MAT and MAP
show significant positive relationships with productivity
\citep{chu_does_2016}. However, as with latitude, the shape of those
relationships is not always clear, and, again, is complicated by the use
of different measures of productivity across studies. Various measures
of primary productivity saturate at high levels of MAP, though the
saturation points identified vary from 1500mm \citep{luyssaert_co_2007}
up to 2445mm MAP \citep{schuur_productivity_2003}. Studies of the
influence of MAT on productivity are less conclusive. Luyssaert et al.
\citeyearpar{luyssaert_co_2007} examined GPP and NPP and found that,
while GPP increases linearly with MAT, NPP saturates at around
10\(^\circ\)C MAT. In contrast, Larjavaara and Muller-Landau
\citeyearpar{larjavaara_temperature_2012}, find that increases in GPP
saturate at approximately 25\(^\circ\)C MAT, while Schuur
\citeyearpar{schuur_productivity_2003} shows that NPP increases linearly
with temperature. bbl:maybe remove following sentences The influence of
these climate variables on productivity is further complicated by the
possibility of interactive effects occurring between them. Taylor et al.
\citeyearpar{taylor_temperature_2017} showed that increased MAP had a
negative effect at low MAT, but a positive effect at high levels of MAT,
and \emph{vice versa} \{\textbf{Helene found this confusing}\},
suggesting that both temperature and water availability are important in
explaining productivity.

\textbf{Primary productivity can be influenced by many other factors as
well}, which often act across a range of scales, and may show
interactive effects with each other
\citep{cleveland_relationships_2011}. On a local scale, stand age
\citep{litton_carbon_2007, gillman_latitude_2015}, biodiversity (Liang
et al. 2019. Positive biodiversity-productivity relationship predominant
in global forests. Science 354, Issue 6309, aaf8957), management
\citep{simova_enigma_2017}; nutrient availability
\citep{aragao_above-_2009}; and altitude
\citep{girardin_net_2010, malhi_variation_2017} all impact forest
productivity. On a global scale, we expect that productivity is most
strongly influenced by broad climatic gradients. There is evidence that
productivity also responds to variables such as cloud cover
\citep{taylor_temperature_2017}, solar radiation
\citep{fyllas_solar_2017}, and potential evapotranspiration
\citep{kerkhoff_plant_2005} in potentially significant ways.
Furthermore, MAT and MAP are very coarse measures of climate, and so
fail to capture much variation in climate on an intra-annual scale,
including the effects of factors such as growing season length, number
of frost-free days, temperature seasonality, and dry season length. Some
studies have suggested that the apparently strong relationship between
MAT and productivity is actually a consequence of the correlation
between MAT and growing season length
\citep{kerkhoff_plant_2005, malhi_productivity_2012, michaletz_convergence_2014, michaletz_drivers_2018}.
Kerkhoff et al. \citeyearpar{kerkhoff_plant_2005} and Michaletz et al.
\citeyearpar{michaletz_convergence_2014} find that, within the growing
season, there is no significant relationship between productivity and
MAT, suggesting that the effect of temperature is due to increased
length of growing season, rather than an inherent influence of
temperature on productivity.

\textbf{The recent development of a global forest carbon database
synthesizing multiple variables and including records of stand history
(ForC;
\citep{anderson-teixeira_carbon_2016, anderson-teixeira_forc:_2018})
opens up the possibility for a standardized analysis of global scale
variation} in multiple components of forest productivity and the
principle climatic drivers of these patterns. In order to approach these
broad and complex issues, we simplify the major gaps in our knowledge to
five broad hypotheses and corresponding specific predictions (Table 1).
** see
\href{https://github.com/forc-db/Global_Productivity/issues/48}{issue
\#48}** First, we ask how forest autotrophic carbon fluxes (FACF) vary
with latitude. We then test how these fluxes relate to MAT and MAP, and
additionally how they respond to other, less well studied, climate
variables. Finally, we consider the relationship between FACF and
seasonality, considering the role of seasonality in explaining variation
in carbon fluxes, and the influence of climate on FACF standardized by
growing season length. We use a comprehensive global database of forest
carbon fluxes to address the above questions for nine carbon fluxes,
allowing for an in-depth exploration of the effect of climate on global
productivity.

\renewcommand{\arraystretch}{2}

\begin{landscape}\begin{table}[!h]

\caption{\label{tab:unnamed-chunk-4}**Summary of hypotheses, corresponding specific predictions, and results.** Direction of significant relationships is indicated by '+' and '-', and  shape is summarized as linear (L), concave up (CU), and concave down (CD). 'n.s.' indicates that the relationship is not signficant.}
\centering
\resizebox{\linewidth}{!}{
\fontsize{12}{14}\selectfont
\begin{tabular}{>{\raggedright\arraybackslash}p{10cm}>{\raggedright\arraybackslash}p{5cm}>{\raggedright\arraybackslash}p{1.8cm}>{\raggedright\arraybackslash}p{1.8cm}>{\raggedright\arraybackslash}p{1.8cm}>{\raggedright\arraybackslash}p{1.8cm}>{\raggedright\arraybackslash}p{1.8cm}>{\raggedright\arraybackslash}p{1.8cm}>{\raggedright\arraybackslash}p{1.8cm}>{\raggedright\arraybackslash}p{1.8cm}>{\raggedright\arraybackslash}p{1.8cm}>{\raggedright\arraybackslash}p{1.8cm}>{\raggedright\arraybackslash}p{1.8cm}}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{1}{c}{ } & \multicolumn{1}{c}{ } & \multicolumn{9}{c}{Forest autotrophic carbon fluxes (FACF)} & \multicolumn{1}{c}{ } \\
\cmidrule(l{3pt}r{3pt}){4-12}
Hypotheses \& Specific Predictions & Related references & Overall & GPP & NPP & ANPP & BNPP & ANPP foliage & ANPP woody stem & BNPP fine root & R auto root & R auto & Support\\
\midrule
\addlinespace[1em]
\hline
\multicolumn{1}{l}{\textbf{H1. FACFs decrease linearly with latitude, and in constant proportion.}}\\
\hspace{1em}1.1. FACFs decrease linearly with latitude (L-) & Luyssaert et al. (2007) & yes & yes & yes & yes & yes & yes & yes & yes & yes & yes & Fig. 2\\
\hspace{1em}1.1.alt. FACFs are similar in tropical and temperate forests, but lower in boreal regions (CD-) & Gillman et al. (2015); Simova and Storch (2017) & no & no & no & no & no & no & no & no & no & no & Fig. 2\\
\hspace{1em}1.2. Allocation of GPP to subsidiary fluxes varies with latitude & Litton et al. (2007); DeLucia et al. (2007) & no & - & no & no & no & no & no & no & - & - & \\
\addlinespace[1em]
\hline
\multicolumn{1}{l}{\textbf{H2. FACFs increase with MAT, with an interactive effect of MAP.}}\\
\hspace{1em}2.1. FACFs increase linearly with MAT (L+) & Schuur (2003) & yes & yes & yes & yes & yes & yes & yes & yes & yes & yes & Fig. 4\\
\hspace{1em}2.2. FACFs increase with precipitation but saturate or decrease at very high levels (CD+) & Luyssaert et al. (2007); Schuur (2003) & yes & yes & yes & yes & yes & yes & n.s. & L+ & L+ & yes & Fig. 4\\
\hspace{1em}2.3. There is a postive interaction between temperature and precipitation (I+) & Taylor et al. (2016) & (yes) & yes & L+ & yes & L- & L+ & yes & L- & yes & yes & Fig. 3\\
\addlinespace[1em]
\hline
\multicolumn{1}{l}{\textbf{H3. FACFs are strongly correlated with other annual climate variables.}}\\
\hspace{1em}3.1. FACFs increases with PET, but saturates or decreases at high levels (CD+) &  & (yes) & yes & yes & yes & yes & L+ & yes & L+ & yes & L+ & Fig. 4\\
\hspace{1em}3.2. FACFs increase with vapour pressure deficit, but saturate or decrease at high levels (CD+) &  & (yes) & yes & yes & yes & yes & yes & yes & L+ & yes & yes & Fig. 4\\
\hspace{1em}3.3. FACFs increase linearly with solar radiation (L+) &  & (yes) & yes & yes & yes & CD+ & yes & CD+ & yes & n.s. & yes & \\
\addlinespace[1em]
\hline
\multicolumn{1}{l}{\textbf{H4. FACFs are reduced under seasonal climates.}}\\
\hspace{1em}4.1. FACFs decrease linearly with temperature seasonality (L-) &  & yes & CU+ & CU+ & CU+ & yes & yes & yes & yes & CU+ & CU+ & Fig. 4\\
\hspace{1em}4.2. FACFs decrease linearly with  precipitation seasonality (L-) &  & no & n.s. & n.s. & n.s. & n.s. & n.s. & n.s. & n.s. & n.s. & n.s. & \\
\hspace{1em}4.3. FACFs increase linearly with growing season length (L+) & Malhi (2012); Michaletz et al. (2014); Chu et al. (2016) & yes & yes & yes & yes & yes & yes & yes & yes & yes & yes & Fig. 4\\
\hspace{1em}4.4. Growing season length is a better predictor of FACFs than MAT & Michaletz et al. (2014); Chu et al. (2016) & no & no & no & no & no & no & no & no & no & no & \\
\addlinespace[1em]
\hline
\multicolumn{1}{l}{\textbf{H5. Considering only growing season months, FACF correlates with climatic drivers.}}\\
\hspace{1em}5.1. Increase with temperature (L+) & Michaletz et al. (2014) & (mixed) & n.s. & n.s. & yes & n.s. & yes & n.s. & n.s. & n.s. & n.s. & \\
\hspace{1em}5.2. Increase with PET (L+) &  & (yes) & yes & yes & n.s. & yes & n.s. & yes & yes & n.s. & n.s. & \\
\hspace{1em}5.3. Increase with precipitation (L+) &  & no & n.s. & n.s. & yes & n.s. & yes & n.s. & n.s. & n.s. & n.s. & \\
\hspace{1em}5.4. Increase with solar radiation (L+) &  & (mixed) & yes & yes & n.s. & yes & n.s. & n.s. & yes & n.s. & n.s. & \\
\bottomrule
\end{tabular}}
\end{table}
\end{landscape}

\subsubsection{Materials and Methods}\label{materials-and-methods}

Analyses were conducted on data contained in the open-access ForC
database
\citep{anderson-teixeira_carbon_2016, anderson-teixeira_forc:_2018}.
This database contains records of field-based measurements of forest
carbon stocks and annual fluxes, compiled from original publications and
existing data compilations and databases. Associated data, such as stand
age, measurement methodologies, and disturbance history, are also
included. The database was significantly expanded since the publication
of \citep{anderson-teixeira_forc:_2018} through integration with the
Global Soil Respiration Database \citep{bond-lamberty_global_2010}.
Additional targeted literature searches were conducted to identify any
further available data on primary productivity, with particular focus on
mature forests in temperate and boreal regions. ForC currently contains
29730 records from 4979 plots, representing 20 distinct ecozones across
all forested biogeographic and climate zones. We used ForC v3.0,
archived on Zenodo with DOI 10.5281/zenodo.3403855.

\begin{figure}[H]
\includegraphics[width=1\linewidth]{distribution_all_variables_cropped} \caption{Map showing all data used in the analysis, coded by variable. Variables are plotted individually in Fig. S1. }\label{fig:unnamed-chunk-5}
\end{figure}

\emph{Data selection.} Over 50 variables of forest carbon stocks and
annual fluxes are represented in the ForC database; this analysis
focused on autotrophic carbon fluxes. Focal variables of this analysis
are listed in Table 2.

A subset of the ForC database was generated for the purposes of this
analysis, in order to control for data quality and remove biasing
factors. Since management can alter observed patterns of primary
productivity \citep{simova_enigma_2017}, sites were excluded from
analysis if they were managed, defined as plots that were planted,
managed as plantations, irrigated, fertilised or including the term
``managed'' in their site description. Sites that had experienced
significant disturbance were also excluded. Disturbances that justified
site exclusion were major cutting or harvesting, and/or burning,
flooding, drought and storm events with site mortality
\textgreater{}10\% of trees. Grazed sites were retained.\\
There is evidence that stand age influences patterns of primary
productivity and carbon allocation in forest ecosystems, and can
confound relationships between latitude and primary productivity
\citep{de_lucia_forest_2007, gillman_latitude_2015}. To reduce any
biasing effects of stand age, we included only stands of known age
\(\ge\) 100 years and those described by terms such as ``mature'',
``intact'', or ``old-growth''.

\newpage

\begin{longtable}[]{@{}llllll@{}}
\caption{Definitions and sample sizes of variables (all annual) used in
analysis. Geographic areas group geographically proximate sites, defined
using a hierarchical cluster analysis on the distance matrix of the
sites, and a cutoff of 25km.}\tabularnewline
\toprule
\begin{minipage}[b]{0.14\columnwidth}\raggedright\strut
Variable\strut
\end{minipage} & \begin{minipage}[b]{0.19\columnwidth}\raggedright\strut
Definition\strut
\end{minipage} & \begin{minipage}[b]{0.13\columnwidth}\raggedright\strut
Components included\strut
\end{minipage} & \begin{minipage}[b]{0.23\columnwidth}\raggedright\strut
Methodologies used\strut
\end{minipage} & \begin{minipage}[b]{0.07\columnwidth}\raggedright\strut
Number of records\strut
\end{minipage} & \begin{minipage}[b]{0.07\columnwidth}\raggedright\strut
Number of geographic areas\strut
\end{minipage}\tabularnewline
\midrule
\endfirsthead
\toprule
\begin{minipage}[b]{0.14\columnwidth}\raggedright\strut
Variable\strut
\end{minipage} & \begin{minipage}[b]{0.19\columnwidth}\raggedright\strut
Definition\strut
\end{minipage} & \begin{minipage}[b]{0.13\columnwidth}\raggedright\strut
Components included\strut
\end{minipage} & \begin{minipage}[b]{0.23\columnwidth}\raggedright\strut
Methodologies used\strut
\end{minipage} & \begin{minipage}[b]{0.07\columnwidth}\raggedright\strut
Number of records\strut
\end{minipage} & \begin{minipage}[b]{0.07\columnwidth}\raggedright\strut
Number of geographic areas\strut
\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
\(GPP\)\strut
\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
Gross Primary Production\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
NA\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
flux partitioning of eddy-covariance; \(NPP\)+\(R_{auto}\)\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
243\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
49\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
\(NPP\)\strut
\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
Net Primary Production\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
stem, foliage, coarse root, fine root, optionally others (e.g., branch,
reproductive, understory)\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
\(ANPP\) + \(BNPP\) (majority); \(GPP\)-\(R_{auto}\)\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
92\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
42\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
\(ANPP\)\strut
\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
Aboveground Net Primary Production\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
stem, foliage, optionally others (e.g., branch, reproductive,
understory) (\textbf{issues
\href{https://github.com/forc-db/Global_Productivity/issues/52}{52},
\href{https://github.com/forc-db/Global_Productivity/issues/54}{54}})\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
\(ANPP_{woody-stem}\) + \(ANPP_{foliage}\) (+ others)\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
256\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
80\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
\(ANPP_{woody-stem}\)\strut
\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
Net Primary Production of woody stems\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
woody stems down to DBH \(\le\) 10cm (no branch turnover)\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
stem growth measurements scaled to biomass using allometries
(\textbf{with recruitment?})\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
264\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
96\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
\(ANPP_{foliage}\)\strut
\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
Net Primary Production of foliage\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
foliage\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
litterfall collection (separated into components)\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
98\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
49\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
\(BNPP_{root}\)\strut
\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
Belowground Net Primary Production\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
coarse and fine roots\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
coarse roots estimated indirectly using allometries based on aboveground
stem increment measures ; fine roots as below\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
101\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
48\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
\(BNPP_{fine.root}\)\strut
\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
Net Primary Production of fine roots\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
fine roots\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
measurements combined one or more of the following: minirhizotrons, fine
root turnover, soil cores, root ingrowth cores\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
88\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
41\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
\(R_{root}\)\strut
\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
Root Respiration\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
coarse and fine roots\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
partitioning of total soil respiration (e.g., through root exclusion),
scaling of root gas exchange\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
64\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
26\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
\(R_{auto}\)\strut
\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
Autotrophic Respiration\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
foliage, stem, and root\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
chamber measurements of component gas exchange\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
22\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
13\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

\emph{Methodological consistency.} The data in ForC is derived from a
range of studies, often employing different methods. For this reason,
criteria were introduced to standardise for differences in methodology.
Where data was based on forest plot census measurements, studies which
used a minimum diameter at breast height (DBH) measure of
\textgreater{}10cm were excluded from analysis. It would be preferable
to standardise by minimum area sampled; however x\% of plots in the
database are 1 ha or under in size; excluding these plots would place
significant constraints on sample size. (\textbf{really, we didn't
exclude any studies based on methodology (all studies measured min DBH
10 or less\ldots{} ) so this paragraph needs revision. Also, address
whether we standardize for inclusion/ exclusion of recruitment, and
address branchfall.})

As discussed above, estimates of \(NPP\), \(ANPP\), and \(BNPP\) are
generated through summing measurements of their component parts. Since
the components included in productivity estimates vary between studies,
estimates of productivity were classified within the ForC database
according to their components, and then filtered for analysis. Estimates
of NPP were selected if they included foliage, branch, stem, coarse
root, and fine root. Measures of NPP which included additional
components, including understorey, volatile organic compounds (VOCs),
exudates, estimates of NPP lost to herbivory (\textbf{these are all
small, no need to exclude}), and the NPP of reproductive structures
(\textbf{often in literfall/ mixed in with foliage\ldots{} did we really
exclude estimates including litterfall?}), were excluded. Estimates of
ANPP were selected if they included foliage, stem growth and optionally
branch turnover. Any measures of primary productivity where components
were unknown were excluded from analysis.

\emph{Climate datasets.} Where site-level data on mean annual
temperature, mean annual precipitation, and latitude were available in
the primary literature, this data was compiled and entered directly into
the ForC database. Based on the geographic co-ordinates for each site,
data on a further 11 climate variables was extracted from five
open-access climate datasets: WorldClim \citep{hijmans_very_2005},
WorldClim2 \citep{fick_worldclim_2017}, the Climate Research Unit (CRU)
time-series dataset v. 4.03 \citep{harris_updated_2014}, the Global
Aridity Index and Potential Evapotranspiration Climate Database
\citep{trabucco_global_2018}, and TerraClimate
\citep{abatzoglou_terraclimate_2018} (see Supplementary Information S1
for details of climate variables). Where site-level data was missing for
mean annual temperature and/or mean annual precipitation, data was
extracted from the WorldClim dataset.

Additionally, two climate variables were derived from the above
datasets: maximum vapour pressure deficit, defined as the vapour
pressure deficit of the month with the largest deficit; and water stress
months, defined as the number of months annually where precipitation was
lower than potential evapotranspiration (PET; \textbf{HML:defined
how?}).

\emph{Length of growing season.} Growing season months were defined as
months with mean minimum temperature \textgreater{} 0.5\(^\circ\)C.
Growing season months were initially calculated following methods used
by Kerkhoff et al. (2005), which additionally required that growing
season months had a moisture index, defined as (MAT - PET)/PET,
\textgreater{} -0.95. Michaletz et al. (2014) included an equivalent
requirement in their calculation of growing season length. However, we
found that including this requirement had no effect on the estimates of
growing season length, and so chose to exclude it.

Monthly data for PET, precipitation, and temperature was downloaded from
the Climate Research Unit (CRU) time-series dataset v 4.03
\citep{harris_updated_2014}, and for solar radiation from WorldClim2
\citep{fick_worldclim_2017}, and used to calculate mean monthly PET,
precipitation, temperature and solar radiation during the growing
season. Total growing season precipitation and solar radiation were also
calculated.

\emph{Model specification.}

\{\textbf{from Helene: It was frequently unclear to me whether the
influences of other climate variables were being considered in
multivariate or univariate analyses. This needs to be clear. I can see
reasons to do it either way, but they need to be differently motivated
and clearly presented.}\}

The effects of climate and latitude on primary productivity were
analysed using mixed effects models using the package `lme4'
\citep{bates_fitting_2015} in R v.3.5.1 \citep{r_core_team_r:_2018}. The
effect of each extracted climate variable on each measure of primary
productivity was modelled by specifying the climate variable as a fixed
effect. For each climate variable, three models were specified: a null
model; a model with the climate variable as a linear term; and a model
with the climate variable as a \textbf{XX}-order polynomial term. AIC
values were calculated for the models and used to select the best model.
If the best model included a polynomial term, the shape of the
polynomial relationship was considered. If the shape of the relationship
\emph{made biological sense}, and was a significant improvement on the
linear relationship (deltaAIC \textgreater{}2), we accepted the
polynomial as the best model. If not, we ran the linear model as the
final model. \(R^2\) values were calculated for the best model. All
\(R^2\) values presented here are marginal \(R^2\) values, and refer to
the proportion of variation explained by only the fixed effects, unless
otherwise specified. In addition, slope coefficients were calculated for
the linear models.

To test for a potential influence of altitude, models were also run with
site altitude included as a second fixed effect. These models were
compared against models with no altitude term, and AIC values calculated
to identify whether inclusion of altitude as a term improved the models.
Including altitude had a very small effect on most models, with the
difference in AIC values between models including and excluding altitude
often being \textless{}2, suggesting the models are very similar in
their explanatory power. As a result, it was decided to present results
only from models without altitude as a fixed term.

Within the ForC database, sites within 25km\textsuperscript{2} of each
other are clustered into geographic areas. To account for correlations
in measurements between tightly clustered sites, a random effect was
specified as plot nested within geographic area. \emph{Data from the
temperate regions was heavily skewed towards studies from the old-growth
forests of the Pacific Northwest.} These forests have very high FACF,
and so to ensure that results were not unduly influenced by geographic
sampling bias, we tried a version of the model where data were weighted
according to forested land area within each Koeppen climate zone.
Results were similar between the weighted and unweighted model. Because
the weighted model may be prone to problems of over-fitting, we present
only results of the unweighted model. \{**Becky, I'd probably just
delete this whole paragraph. Since we ended up not using the weighted
model, it's just extra info for readers to digest.\}

Models were run for total annual FACF against annual climate variables,
and for fluxes standardized by growing season length, defined as annual
FACF/length of growing season, against growing season climate variables.
For analyses on data within the growing season, only linear models were
specified.

\{\textbf{Note from Helene on the following paragraph: this makes it
sound like analyses from H3 were about variation remaining after
accounting for temperature, but analyses below don't seem to do
that.}\}\\
To investigate the potential interactive effects of climate variables on
carbon fluxes, multivariate models were also specified. To ensure that
models were biologically meaningful, the terms included in the models
tested built on results from the univariate models. Modelling of
individual climate variables identified that the best predictors of
carbon fluxes were variables related to temperature. We therefore
decided to include mean annual temperature as a term in all multivariate
models. We first modelled the interaction effect between mean annual
temperature and mean annual precipitation, in order to capture climate
variation along the axes of temperature and water availability. Models
were tested for a significant interactive effect and a significant
additive effect. We then explored whether any other climate variable, in
combination with mean annual temperature, could significantly improve on
the combination of mean annual temperature and mean annual
precipitation. In specifying the range of models to test, climate
variables which were strongly correlated with temperature were dropped,
in order to capture the greatest range of variation in climate. For each
possible pairing of climate variables, two models were specified: a
model with the two climate variables showing an additive effect; and a
model with the two climate variables showing an interactive effect. As
before, plot nested within geographic area was included as a random
effect. Altitude was not considered. AIC values were calculated for the
models, and used to compare models. Models were considered to be
significantly better than the baseline MAT*MAP model if:\\
i) the AIC value of the model was smaller than the AIC value of the
baseline model by \textgreater{}2\\
ii) the r-squared value of the model was larger than the r-squared value
of the baseline model by \textbf{\textgreater{}5 (???- not possible)}

\emph{Validating models of component fluxes.} Comparison of component
fluxes is based on the assumption that components sum accurately to
estimates of larger fluxes. To test this, components of larger fluxes
were regressed against latitude, and the models used to generate a
series of point estimates along lines of best fit for each component.
The point estimates for smaller component fluxes were summed to generate
new ``stacked'' estimates of larger fluxes, which were then compared
against actual measurements of the larger flux. Confidence intervals for
the larger flux were calculated using the `bootMer' function from the
lme4 package \citep{bates_fitting_2015}. Stacked plots were generated
for:\\
1. GPP = NPP + R\textsubscript{auto}\\
2. NPP = ANPP + BNPP\\
3. ANPP = ANPP\textsubscript{foliage} + ANPP\textsubscript{woody}
\textsubscript{stem}\\
4. Total belowground carbon flux = BNPP + R\textsubscript{root}

\emph{Allocation to carbon fluxes along latitudinal gradients.}
Variation in allocation to component carbon fluxes along latitudinal
gradients was explored for a range of pairings: firstly, GPP:NPP,
ANPP:BNPP, and ANPP\textsubscript{foliage}:ANPP\textsubscript{woody}
\textsubscript{stem}; and secondly, the ratio of NPP to each of ANPP,
BNPP, ANPP\textsubscript{foliage}, and ANPP\textsubscript{woody}
\textsubscript{stem}. For each set of paired fluxes, measurements taken
at the same site and plot, and in the same year, were paired together,
and the ratio of each pair of measurements calculated. The ratios were
regressed against latitude and climate variables, using the linear model
specified above. Cook's distance analyses were carried out for each of
the models, and indicated that data from a few high-elevation sites were
having a disproportionate influence on the regressions. To account for
this, models were re-run using only data from sites \(\le\) 1000m.

\subsubsection{Results}\label{results}

In total, we analyzed 1228 records from 9 C flux variables taken from
forests that had experienced no major anthropogenic disturbances within
the past 100 years. These records represented a total of 154 distinct
geographic areas (Fig. 1, Table 2), across all forested biogeographic
and climate zones.

\emph{How does productivity vary with latitude?}

All major carbon fluxes decreased linearly with latitude (Fig. 2).

Latitude was a strong predictor for many of the carbon fluxes,
explaining 64\% of variation in GPP (n = 254, p\textless{}0.0001), 50\%
in NPP (n = 114, p\textless{}0.0001) and 45\% in ANPP (n = 259,
p\textless{}0.0001). For all fluxes, their relationship with latitude
was best predicted by the linear model.

\emph{Relationships and differences among fluxes.} In general, smaller
component fluxes summed approximately to larger fluxes across the
latitudinal gradient (fig. 2). That is, modelled estimates of GPP,
generated from the sum of NPP and R auto; NPP, generated from the sum of
ANPP and BNPP\textsubscript{root}; and ANPP, generated from the sum of
ANPP\textsubscript{foliage} and ANPP\textsubscript{woody}
\textsubscript{stem}, fell completely within the confidence intervals of
the regressions of field estimates of GPP, NPP and ANPP respectively.

bbl: here or in the discussion, note that this is a fairly stringent
test: it's easy for sub-fluxes not to sum up! (Extensive citations from
EC literature for example.). HML: yes and no. confidence intervals are
pretty large.

\begin{figure}[H]
\includegraphics[width=1\linewidth]{combined_stacked} \caption{Latitudinal trends in forest autotropic carbon flux. Lines of best fit are plotted according to the best model selected during analysis. All regressions are significant $(p<0.05)$. Each panel shows major C fluxes together with component fluxes. Also plotted are predicted trends in the major C fluxes based on the sum of component fluxes. 95\% confidence intervals are plotted for the major flux for comparison with predicted trends. In (d),  which shows three belowground fluxes, the major flux, total belowground carbon flux, is one for which we have no data}\label{fig:unnamed-chunk-6}
\end{figure}

We found no evidence that allocation between fluxes varied substantially
with latitude or climate. There were no significant results from
regressing ratios of carbon fluxes against latitude, or against any of
the climate variables.

\(R^2\) values were generally highest in the major fluxes, and decreased
in subsidiary fluxes (Supporting Information S2). Of the major fluxes,
R\textsubscript{auto} and GPP were the most strongly explained by
latitude and climate, with climate explaining at most 71\% of variation
in GPP, and 65\% in R\textsubscript{auto}. The proportion of variation
explained by climate and latitude decreased in NPP and ANPP, with
climate explaining at most 51\% of variation in NPP and 44\% in ANPP. Of
the major fluxes, BNPP\textsubscript{root} was the least well explained
by climate and latitude, with climate explaining at most 36\% of
variation.

With the exception of ANPP\textsubscript{foliage}, the proportion of
variation explained by climate and latitude in subsidiary fluxes was
much lower. Climate explained at most 24\% of the variation in
ANPP\textsubscript{woody} \textsubscript{stem}, 19\% in
BNPP\textsubscript{fine} \textsubscript{root}, and 27\% in
R\textsubscript{root}. In contrast, climate strongly explained variation
in ANPP\textsubscript{foliage}, with mean annual temperature explaining
58\% of variation. This pattern was also seen in the \(R^2\) values for
multivariate models.

\emph{How does productivity relate to MAT and MAP?} MAT and MAP are the
most commonly reported site-level climate variables, and much previous
research into the effect of climate on forest productivity has focused
on these as key climate variables. MAT was a significant
(p\textless{}0.05) and strong predictor of productivity for all carbon
fluxes tested, with all fluxes showing a linear increase with
temperature (fig. 4). We found no support for a saturation point of
productivity with temperature.

MAP was found to be a significant (p\textless{}0.05) but poor predictor
of productivity, explaining, with the exception of
R\textsubscript{auto}, at most 37\% of variation in carbon flux. For the
majority of fluxes productivity was best predicted by a polynomial
model. Productivity increased with precipitation, up until a saturation
point at between 3000 and 4000mm annual precipitation, above which
productivity started to decrease (fig. 4). The notable exception to this
was GPP: the model indicated that GPP continued to increase with
precipitation up to measures of at least 5000mm annually
(p\textless{}0.0001, \(R^2\) = 0.33. Data above this point was not
available, but the model trend indicated that the saturation point for
this model would be around 5000mm MAP.

There was a significant interactive effect between MAT and MAP for GPP,
BNPP\textsubscript{root}, BNPP\textsubscript{fine} \textsubscript{root},
ANPP, ANPP\textsubscript{woody} \textsubscript{stem}, and
R\textsubscript{root} (fig.3). There was a significant additive effect
for R\textsubscript{auto}. NPP and ANPP\textsubscript{foliage} showed no
significant interactive or additive effect. \{\textbf{NEED SI TABLE
GIVING FITTED COEFFICIENTS \& STATS}\}

For the variables which showed a significant interactive or additive
effect between MAT and MAP, no other climate variable, in combination
with MAT, significantly improved on that model. For NPP, there was a
significant interactive effect between MAT and water stress months, with
this model explaining nearly 5\% more variation in NPP than MAT alone
\{\textbf{Helene: but less than MATxMAP? if so, why bother
mentioning?}\}. However, for ANPP\textsubscript{foliage}, no
multivariate model improved on the univariate model including only MAT.

\begin{figure}[H]
\includegraphics[width=1\linewidth]{mat_map_interaction} \caption{Interactive effects of mean annual temperature and mean annual precipitation on FACF. For visualization purposes, data points are grouped into bins of 0 - 1000, 1001 - 2000, 2001 - 3000, and >3000mm mean annual precipitation, and lines of best fit models are plotted for mean annual precipitation values of 500, 1500, 2500, and 3500mm. All regressions are significant $(p<0.05)$.}\label{fig:unnamed-chunk-7}
\end{figure}

\emph{How does productivity relate to other climate variables?}
\{\textbf{Here, maybe add some statement about covariation among climate
variables, site SI figure}\} Our results indicated that productivity was
most strongly explained by temperature at the global scale, with
temperature-related climate variables coming out as strong predictors of
productivity. In addition to MAT, temperature seasonality, annual
temperature range, and annual frost days were consistently identified as
strong univariate predictors of FACF. \{\textbf{cite some table/ SI?}\}

We found a significant relationship between productivity and potential
evapotranspiration for all fluxes. ANPP\textsubscript{foliage},
BNPP\textsubscript{fine} \textsubscript{root} and R\textsubscript{root}
increased linearly with PET; however, all other fluxes showed a
polynomial relationship with PET (fig. 4). We found strong evidence for
a saturation point or peak with PET: productivity tended to increase at
values below 1000mm, before saturating between 1200 and 1700mm. There
was also evidence that productivity begins to decrease at values above
1800mm PET.

Vapour pressure deficit was a significant predictor of productivity for
all fluxes. BNPP\textsubscript{fine} \textsubscript{root} showed a
linear relationship with vapour pressure deficit (\(R^2\) = 0.07,
p\textless{}0.05), but all other fluxes showed a polynomial relationship
(fig. 4). Productivity initially increased with vapour pressure deficit,
before saturating at around 0.8 kPa. At values above 0.8 kPa,
productivity began to decrease.

All fluxes, with the exception of R\textsubscript{root}, showed a
positive linear relationship with solar radiation. Solar radiation
explained a low proportion of variability in productivity for all
fluxes, explaining less than 20\% of the variation in each flux, with
the exception of R\textsubscript{auto} (\(R^2\) = 0.26,
p\textless{}0.05).

Of the climate variables tested, annual wet days, aridity, cloud cover,
mean diurnal temperature range, precipitation seasonality, maximum
vapour pressure deficit and water stress months were poor or
non-significant explainers of variation in productivity, explaining less
than 20\% of the variation in each of the carbon fluxes.

\begin{figure}[H]
\includegraphics[width=1\linewidth]{combined_plots} \caption{Plots of carbon fluxes against (a) mean annual temperature; (b) mean annual precipitation; (c) potential evapotranspiration, (d) vapour pressure deficit; (e) temperature seasonality; (f) length of growing season. For visualization purposes, data for each flux was rescaled with a mean of 0 and standard deviation of 1. Lines of best fit are plotted according to the best model selected during analysis (**see issue 47**). All regressions are significant $(p<0.05)$.}\label{fig:unnamed-chunk-8}
\end{figure}

\emph{What is the role of seasonality in explaining FACF?} Temperature
seasonality was a significant predictor of FACF. GPP, NPP, ANPP, and
R\textsubscript{root} exhibited a polynomial relationship with
seasonality (Fig. 4) \{\textbf{give statistic on signficiance}\}.
ANPP\textsubscript{foliage}, ANPP\textsubscript{woody}
\textsubscript{stem} and R\textsubscript{auto} decreased linearly with
temperature seasonality. Temperature seasonality was strongly correlated
with annual temperature range, and, as expected, all fluxes showed
almost identical responses to it \{\textbf{meaning unclear}\}.
Productivity was highest where temperature seasonality = 0, and at an
annual temperature range of 15\(^\circ\)C or lower. In contrast, there
was no significant effect of precipitation seasonality on FACF.

We found a significant relationship between length of growing season and
FACF, with all fluxes showing a linear increase in productivity with
length of growing season (Fig. 4). Length of growing season was a strong
predictor of FACF, explaining 51\% of variation in GPP, 39\% of
variation in NPP, and 34\% of variation in ANPP, but it was a weaker
predictor than MAT for all fluxes analysed.

\emph{Within the growing season, how does FACF vary with climate?} When
FACF was standardized by growing season length, correlations with
climate variables were much weaker. Growing season temperature,
precipitation, solar radiation, and PET each had a small effect on one
or more of the nine FACF analyzed (Table 2; Fig. \textbf{S\#}).
Speficifally, \(ANPP\) increased with growing season temperature
(\(R^2\) = 0.10, p\textless{}0.001) and precipitation (\(R^2\) = 0.04,
p\textless{}0.05). Similaryly, \(ANPP_{foliage}\) increased slightly
with growing season temperature (\(R^2\) = 0.16, p\textless{}0.01) and
precipitation (\(R^2\) = 0.09, p\textless{}0.05). Growing season solar
radiation had a positive influence on \(GPP\) (\(R^2\) = 0.21,
p\textless{}0.001), \(NPP\) (\(R^2\) = 0.21, p\textless{}0.001),
\(BNPP\) (\(R^2\) = 0.16, p\textless{}0.001) and \(BNPP_{fine.root}\)
(\(R^2\) = 0.12, p\textless{}0.01). Growing season PET had a positive
influence on \(GPP\) (\(R^2\) = 0.15, p\textless{}0.01), \(NPP\)
(\(R^2\) = 0.18, p\textless{}0.01), \(BNPP\) (\(R^2\) = 0.23,
p\textless{}0.0001), \(BNPP_{fine.root}\) (\(R^2\) = 0.11,
p\textless{}0.05), and \(ANPP_{woody-stem}\) (\(R^2\) = 0.06,
p\textless{}0.05). \{\textbf{Becky, please verify/ edit the following:}
There were no other signifcant correlations between growing season
length-standardized FACF (9 variables in Table 2) and growing season
climate (\textbf{which variables?})\}.

\subsubsection{Discussion}\label{discussion}

We used a large global database (ForC), containing an unprecedented
amount of data representing all major forest types (Fig. 1) and the nine
most significant forest autotrophic carbon flux (FACF) variables (Table
2), to comprehensively explore how C cycling in mature, undisturbed
forests varies with latitude and climate on a global scale. We test a
suite of hypotheses, including many previously posed (Table 1),
standardising for factors that have not always been controlled for
(\emph{e.g.}, stand age, flux components measured) to gain a stronger
understanding of climatic effects on FACF.

bbl: Most of these two paragraphs (following) just restate the results.
Honestly I would remove them enitrely, but at the very least, condense
into a single paragraph.

We show that, across all nine FACF variables analyzed, C cycling
decreases linearly with latitude (\emph{H1.1}; Fig. 2)--a finding that
confirms multiple previous studies (\textbf{REFS}) but contradicts the
idea that productivity of temperate forests rivals that of tropical
forests (\emph{H1.1.alt}; \textbf{REFS}). The FACF variables increase in
proportion to one another (\emph{H1.2}), with no differences in
allocation detectable at this global scale and with component fluxes
summing appropriately to larger fluxes (Fig. 2), indicating no major,
systematic omissions or overestimations of flux components. However, we
did detect a tendency for greater variability among subsidiary C fluxes
(e.g., \emph{\(ANPP_{woody}\)}, \emph{\(BNPP_{fine.root}\)}; Fig. 2;
\textbf{some SI table?}).

\textbf{Climate explains a significant proportion of variation in all C
fluxes analyzed} (albiet less for subsidiary fluxes), with temperature
variables the best predictors of FACF at this global scale (Figs. 3-4).
While other climate variables are significant predictors of FACF
(\emph{H3}), none of them in isolation improve on the explanatory power
of temperature-related variables in general or MAT specifically (Fig.
4). Water availability is an important factor in explaining FACF on a
global scale: we find a positive influence of precipitation at low MAP,
with saturation at higher levels of MAP (Fig. 4b) and a significant
interaction between MAT and MAP for seven of the nine variables analyzed
(Fig. 3). Temperature seasonality and growing season length are closely
correlated with MAT and are strong predictors of FACF (\emph{H4}; Fig.
4e-f), though growing season length doesn't improve upon MAT as a
predictor. Within the growing season, the influence of climate on C
cycling is smaller but still significant for a number of carbon fluxes
(\emph{H5}; \textbf{some SI table?}). These findings clarify the big
picture of how FACF varies with latitude and climate on a global scale.

Past studies have differed in their conclusions regarding the
relationship between productivity and latitude or its correlates (Table
1, \emph{H1}; \textbf{REFS})--quite possibly because of lack of
standardization with respect to stand age and disturbance history. Our
findings indicate that, among mature, undisturbed stands, FACF is
unambiguously highest in the tropical regions, and the relationship is
approximately linear (Fig. 2). This contrasts with suggestions that
productivity of temperate forests is similar to that of tropical forests
(\textbf{REFS}). Temperate forests tend to be younger than tropical
forests (\textbf{REF}), so analyses comparing across latitudinal
gradients without controlling for stand age risk confounding age with
biome effects (\href{http://pure.iiasa.ac.at/id/eprint/15361/}{Poulter
et al. 2018-DOI:10.1594/PANGAEA.889943}; Potapov et al. 2008. Mapping
the World's Intact Forest Landscapes by Remote Sensing. Ecology and
Society 13 (2), 51. ). In addition, because C allocation varies with
stand age \citep{de_lucia_forest_2007} (\textbf{See Nobby's comment in
manuscript-draft\_NK.pdf}), age differences may introduce systematic
biases into analyses of FACF across latitude or global climatic
gradients. For example, woody productivity tends to be higher in rapidly
aggrading secondary stands than in old-growth forests, where
proportionally more C is allocated to respiration (** Nobby AFM paper:
Understanding the controls over forest carbon use efficiency on small
spatial scales: Effects of forest disturbance and tree diversity**)
{[}*purpose for respiration/ other compenents (\textbf{See Nobby's
comment in manuscript-draft\_NK.pdf}){]}.

We show that FACF are broadly consistent in their responses to climate
drivers on the global scale (with the exception of some differences in
MAT-MAP interactions; Fig. 3), with no major trends in C allocation
among the variable pairs tested (Table 1, \emph{H1}; Fig. 2;
\textbf{Some SI table}). Although variation in allocation has been
observed along gradients of elevation \citep{moser_elevation_2011} and
water availability \citep{newman_above-_2006}--along with non-climatic
axes of stand age \citep{litton_carbon_2007}, nutrient availability
\citep{litton_carbon_2007, gill_belowground_2016}, and forest structure
\citep{taylor_greater_2019}--variation in relation to climate is not
apparent at the global scale within ForC, which contains the bulk of
relevant data. Our conclusion, then, is that hypothesized gradients in
allocation along global climate gradients cannot currently be supported
for mature forests, although data quantity and standardization is
currently insufficent to rule out the possibility that such trends
exist.

bbl: remove this sentence? Of particular interest and significance are
the relationships amongst \(GPP\), net primary productivity (\(NPP\) and
its components, particularly \(ANPP_{woody-stem}\)), and respiration
(\(R_{auto}\) and components). There have been suggestions that tropical
forests tend to have low carbon use efficiency (\(CUE\)=
\(NPP\)/\(GPP\)=(\(GPP\)-\(R_{auto}\)/\(GPP\)), which are based on
observations of low \(CUE\) in old-growth tropical forests relative to
(mostly younger) extratropical forests
\citetext{\citealp{de_lucia_forest_2007}; \citealp[\href{https://www.researchgate.net/publication/227837530_The_productivity_metabolism_and_carbon_cycle_of_tropical_forest_vegetation}{Malhi
2012};][]{anderson-teixeira_carbon_2016}}, but our analysis suggests
that these low values might more appropriately be attributed to the fact
that these forests are old than to their tropical climate. Indeed,
\(CUE\) is known to decline with forest age
(\citep{de_lucia_forest_2007};
\href{https://academic.oup.com/treephys/article-abstract/39/8/1473/5423351?redirectedFrom=fulltext}{Collalti
\& Prentice 2019}) but appears to be roughly independent of \(GPP\)
\citep{litton_carbon_2007}. Among our sites with relevant data, there is
indication that CUE or \(ANPP_{woody-stem}\)/\(GPP\) increase with
latitude (\textbf{some SI table}). Additional measurements with careful
methodological standardization across a consistent set of mature forest
sites will be necessary to determine whether any climate-driven
gradients in allocation exist at the global scale.

One interesting observation was that climate tends to explain more
variation in the major fluxes (\(GPP\), \(NPP\), \(R_{auto}\) - latitude
\(R^2\ge\) 48\%) than in subsidiary fluxes (latitude \(R^2\)
\textless{}27\% for \(BNPP_{fine.root}\),
\(R_{auto-root}\),\(ANPP_{woody-stem}\)) (Fig. 2; \textbf{some SI
table?}). There are two, non-exclusive, potential explanations for this.
First, it may be that methodological variation is larger relative to
flux magnitude for some of the subsidiary fluxes. Belowground fluxes in
particular are difficult to quantify, and measurement methods for the
belowground fluxes considered here may be measured through fundamentally
different approaches (\emph{e.g.}, minirhizotrons, ingrowth cores, or
sequential coring for \(BNPP_{root-fine}\); root exclusion, stable
isotope tracking, or gas exchange of excised roots for
\(R_{auto-root}\)), and sampling depth is variable and often
insufficient to capture the full soil profile. \(ANPP_{woody-stem}\),
which is also poorly explained by latitude or climate, is more
straightforward to measure but is subject to variability introduced by
differences such as biomass allometries applied and minimum plant size
sampled (bbl: cite e.g.~Huntzinger?). However, methodological variation
and uncertainty affect all of fluxes considered here--not necessarily
any less than the aforementioned, and some of the larger fluxes that
vary more strongly with respect to climate (\(ANPP\), \(NPP\)) are
estimated by summing uncertain component fluxes. Second, differences
among variables in the proportion of variation explained by climate may
be attributable to more dicrect climatic control over \(GPP\) than
subsidiary fluxes. That is, subsidiary fluxes may be shaped by climate
both indirectly through its influence on \(GPP\) and respiration and
directly through any climatic influence on C allocation, as well as many
other local- and regional-scale factors (\textbf{REFS}).

The latitudinal gradient in FACF (Fig. 2) is driven primarily by
temperature-related climate variables, and secondarily by moisture
availability (Table 1, \emph{H2-H3}; Figs. 3-4). Because many climate
variables co-vary across the latitudinal gradient, because climatic
drivers affect forest carbon flux on much shorter time scales than can
be captured by annual climate summary variables, and because both
climatic conditions and C flux vary intra- and inter-annually around the
long-term means, it is not appropriate to attempt to identify any one
mean annual climate variable as a mechanistic driver of FACF. However,
it remains informative to consider these relationships. We find that
temperature-related climate variables (\(MAT\), temperature seasonality,
\ldots{}\textbf{LIST}) explain the highest proportion of variability in
productivity, and among these, \(MAT\) is generally the best
predictor--perhaps because site-specific MAT is recorded for the
majority of sites in ForC, whereas other variables are extracted from
global gridded data products (Table S1\textbf{?}). The effects of
temperature are modified by moisture availability, with reduced FACF
under hot and dry conditions (\emph{i.e.}, high PET, high deficit; Fig.
4c-d) and sometimes under very high precipitaiton (Figs. 3, 4b).
Negative effects of very high precipitation on FACF have been previously
observed (\textbf{REFS}) and are attributable to nutrient and light
limitations (\textbf{REFS}) Thus, although temperature and water
availability jointly and interactively drive global-scale patterns of
FACF.

bbl: this seems really interesting and novel, consider expanding the
following paragraph

FACF is negatively correlated with temperature seasonality (Table 1,
\emph{H4}; Fig. 4e), and productivity is shut down during cold- or dry-
dormant seasons. To account for this, a number of analyses seeking to
characterize global-scale effects of climate on productivity have
examined the relationship of FACF per month of the growing season with
growing season climatic conditions (Table 1, \emph{H5}; \textbf{REFS}).
We found that the sort of simple metric needed to define growing season
at a global scale was uncertain for temperature and problematic for
moisture (\textbf{WORK ON THIS}). A temperature-defined growing season
length had stong positive correlation with FACF (Fig. 4f), but explained
less variation than \(MAT\). Dividing FACFs by growing season length to
yield FACF per growing season month removed the majority of
climate-related variation, supporting the idea that the latitudinal
gradient in FACF is attributable more to shorter growing seasons at high
latitudes than to inherently lower rates of photosyntheiss or
respiration by high-latitude forests (\emph{{[} Enquist et al. 2007 GCB-
but check{]}}). However, there remained a number of significant
correlations with growing season climatic conditions, suggesting that
climatic conditions remain influential within the growing season. We
conclude that while correcting for growing season length takes analyses
a step closer to mechanistic linkage of instantaneous C flux rates to
environmental conditions, it remains very crude relative to the the
timescales on which climate affects plant metabolism and does not
advance statistical predictive power. Rather, mechanistic accounting for
climatic effects on global productivity patterns requires models
representing physiologically meaningful timescales (\emph{e.g.},
\textbf{refs}).

Our analysis clarifies how FACF varies (bbl: is FACF singular or plural?
Be consistent) with latitude and climate on a global scale, with some
important implications for how forest carbon cycling relates to climate
and, by extension, how it is likely to respond to climatic warming. We
find no support for non-linear trends in mature forest C cycling with
respect to latitude or \(MAT\), and no distinct trends in C allocation
across the global scale (Fig. 2). The implication is that under warmer
conditions with similar moisture availability--and within the
temperature range to which forest communities are adapated and
acclimatized--higher temperatures result in a generalized acceleration
of FACF, with no major shifts in C allocation among subsidiary fluxes.
Of course, actual climatic changes will result in very different sets of
conditions than represented across geographic gradients in climate, but
our analysis clarifies how carbon cycles through contemporary forest
ecosystems. As we enter a period of accelerating climatic change,
understanding of the fundamental climatic controls on FACF sets a
foundation for understanding patterns of change. \textbf{{[}work on
this{]}}

\subsubsection{Acknowledgements}\label{acknowledgements}

Scholarly Studies ForestGEO Compilation of the ForC database was
originally funded by DOE

\bibliography{library.bib,packages.bib}

\end{document}