### 2020-08-06 manuscript/journal_submissions/2020_8_GCB/submission_notes.Rmd
add submission notes

---
title: "GCB submission notes"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# submission questions

## What is the scientific question you are addressing?
## What is/are the key finding(s) that answers this question?
##    Why is this work important and timely?
##    Does your paper fall within the scope of GCB; what biological AND global change aspects does it address?
##    What are the three most recently published papers that are relevant to this question? This information will assist the Editors in selecting reviewers.